Data-driven approaches to identifying knowledge gaps in marine invertebrate biodiversity

### 2017-05-19 manuscript.Rmd
more edits to manuscript

With increasing impact of human activities on the biosphere, we need efficient and accurate approaches to document changes in species ranges and community composition. While terrestrial ecosystems have received considerable attention, the baseline of healthy and resilient marine ecosystems are not well understood, and modeling the impact of human activities is limited to a few taxonomic groups. A major hinderance is that we do not have enough information about which species live where, and an unknown but potentially significant portion of marine biodiversity remains to be discovered and described. Additionally, marine ecosystem monitoring initiatives are lacking to detect and understand, how and why communities change through time [@Duffy2013,@Borja2016]. There is therefore a pressing need to improve our knowledge of marine systems before species go extinct, and their functionning are irreversibly altered. Our lack of knowledge about marine invertebrates make the modelisation of trophic networks less accurate, as diverse species assemblages are often reduced as simple functional categories that do no account for their ecological diversity. Until very recently, our knowledge of biodiversity was scattered among heterogeneous sources (e.g. natural history collections, publications, expert knowledge), and difficult to access. This heterogeneity and the difficulty to access this data have made synthesis and the identification of knowledge gaps difficult. However, with the development of linked, public, comprehensive databases of natural history records (e.g. iDigBio), species occurrences (e.g., GBIF, OBIS), genetic data (e.g., GenBank, BOLD), and registers of species names (e.g., WoRMS), it is now easier to quantitavely evaluate the state of our knowledge for biodiversity data. Here, we aggregate publicly available data for marine invertebrates that occur in the United States to highligth gaps in our knowledge. Given the urgency of the biodiversity crisis, we need to develop data-driven initiatives to efficiently document biodiversity, and build on our existing knowledge to collect information to improve our understanding of marine systems. Additionally, we emphasize that our understanding of marine biodiversity, and the impact of human activities on communities require not only to get a more accurate list of the species that inhabits the oceans, but we also need better data on species distributions and abundances. This can only be achieved with additional sampling and a commitment to training more taxonomists and experts that can identify the material collected. We live at a time when it has never been easier and cost effective to survey comprehensively the extant of marine biodiversity.

Biodiversity discovery and documentation has been transformed in the 21st century by using molecular data to facilitate species identification, delineate species. And more recently, the development of high throughput sequencing technologies is used to quantify diversity and uncover large scale patterns of biodiversity organization. The wide adoption of DNA barcoding (where homologous DNA sequences are sequenced across the main domains of the Eukaryotes) has unraveled the pervasiveness of cryptic species, invigorating taxonomy and the description of new species, and enabled collaborative research programs such as the Census of Marine Life. Unraveling these cryptic species has also contributed to the recognition that species have smaller ranges than their taxonomic identity had implied making them potentially more susceptible to extinction or overexploitation. Despite early criticisms and now well-understood limitations, DNA barcoding has been suggested as a possible solution to the widening gap between the increased need to document biodiversity while taxonomic expertise is dwindling. Thus, large databases of DNA barcodes have been collaboratively assembled and made pubicly available. By matching DNA barcodes of unidentified species to a reference public database of accurately identified and vouchered specimens, DNA barcoding can greatly speed up species identification. Sequences with no match make it easier to flag specimens that warrant further taxonomic attention.

Recent developments in high-throughput sequencing technologies are giving new applications to these DNA barcode databases [@Leray2016]. Metabarcoding has been developed to sequence in parallel the barcoding gene across the species included in environmental samples such as plankton tows, fish guts, or artificial structure scrappings [@Leray2015]. With decreasing sequencing costs, and the prospect of automating sample collection, these technologies will soon be used routinely for ecosystem and biodiversity monitoring [@Bohan2017]. DNA barcoding provides the possibility to link these newly generated data with historical natural history knowledge about species and their distributions. However, the insights we will gain from these approaches are only going to be as good as the level of completeness and accuracy of the information included in reference databases. However, to accurately identify the species found in these samples, comprehensive DNA barcode databases are needed. Here, we mostly focus on museum specimens because we can validate their identification, and a fraction of these samples (if they have not been preserved in formalin) can be used to obtain DNA from them to complete existing databases of DNA barcodes.

Our knowledge about species and their distributions is gathered in regional faunistic lists, typically assembled by taxonomic experts. They are comprehensive but limited in their geographic scope. On the other hand, natural history collections have more comprehensive geographic coverage, but the quality of the taxonomic identifications can vary, and some common species might be underrepresented. We assessed the level of completeness of biodiversity public databases for marine invertebrates that occur within the boundaries of the EEZ of the US. Specifically, we compared comprehensive regional faunistic lists to assess how much of this knowledge is captured by current digitization efforts of museum collections and occurrence records; and we estimate how many of the species have genetic data publicly available that would be useful in the context of identifying samples from metabarcoding studies.

@@

@@

@@

>>>>

With increasing impact of human activities on the biosphere, we need efficient and accurate approaches to document changes in species ranges and community composition. While terrestrial ecosystems have received considerable attention, the baseline of healthy and resilient marine ecosystems are not well understood, and modeling the impact of human activities is limited to a few taxonomic groups. A major hinderance is that we do not have enough information about which species live where, and an unknown but potentially significant portion of marine biodiversity remains to be discovered and described. Additionally, coordinated marine ecosystem monitoring initiatives are lacking [@Duffy2013,@Borja2016]. There is therefore a pressing need to improve our knowledge of marine systems before species go extinct, and their functionning are irreversibly altered.

Until very recently, our knowledge of biodiversity was scattered among heterogeneous sources (e.g. natural history collections, publications, expert knowledge), and difficult to access. These factors have contributed to the lack of synthesis and have made the identification of knowledge gaps difficult. However, with the development of public databases for natural history records (e.g. iDigBio [@Page2015]), species occurrences (e.g., GBIF [@Edwards2004], OBIS [@Zhang2002]), genetic data (e.g., GenBank [@Benson2013], BOLD [@Ratnasingham2007a]), and registers of species names (e.g., WoRMS [@Costello2013b]), we can quantitavely evaluate what we know. Here, we aggregate publicly available data for marine invertebrates that occur in the United States and in the Gulf of Mexico to highlight gaps in our knowledge. We show that public databases are relatively incomplete, and that most species are known from a few speciemens and a few locations. We emphasize that our understanding of marine biodiversity, and the impact of human activities on communities require not only to get a more accurate list of the species that inhabits the oceans, but we also need more comprehensive data on species distributions and abundances. This can only be achieved with additional sampling and a commitment to training more taxonomists and experts that can identify the material collected.

Regional faunistic lists assembled and curated by taxonomic experts typically provide a comprehensive and detailed account of species that are known to occur in a given geographical area. We assessed how faunistic lists assembled over decades by specialists compare with records in public databases. We parsed faunistic lists for marine invertebrates from the Gulf of Mexico [@Tunnell2009] and from the Pacific Northwest [@Kozloff1987]. Our goals were (1) to compare  the species lists for these geographical areas assembled by experts with the records found in databases; (2) to assess the proportion of species included in the expert lists that have records in public databases.

To compare the faunistic lists from the Gulf of Mexico and the Pacific Northwest with records in public databases, we queried the iDigBio and OBIS databases for all records they included in the same geographic boundaries as the lists. After filtering these lists to only include the same higher taxonomic level taxa as the species lists, we had for the Gulf of Mexico `r nrow(idigbio_gom_records)` and `r nrow(obis_gom_records)` records for iDigBio and OBIS respectively representing `r dplyr::n_distinct(idigbio_gom_records, worms_valid_name)` and `r dplyr::n_distinct(obis_gom_records, worms_valid_name)` species. For the Gulf of Mexico, most of the species included in the expert lists are recorded in both iDigBio and OBIS (xx %), or OBIS (xx %). However, for the Pacific Northwest, only xx% of the species from the expert list have records in public databases.

assess the proportion of species included in these lists that (1) are recorded in databases from the same geographical area; (2) have any records in databases
, with the records that occur in the same geographical areas in iDigBio and OBIS. Our goal was to assess the proportion of species known to occur in a given area by specialists that have been captured by digitization efforts. First, to estimate the proportion of species included in these faunistic lists whose occurences have been digitized, we queried biodiversity databases using the same geographical limits as the ones used by the authors of these faunistic lists. We found ... Second, to estimate whether listed species are known to at least occur in these databases, we queried them for the presence of the species within the databases.

@@

; and we estimate how many of the species have genetic data publicly available that would be useful in the context of identifying samples from metabarcoding studies

@@

Given the urgency of the biodiversity crisis, we need to develop data-driven initiatives to efficiently document biodiversity, and build on our existing knowledge to collect information to improve our understanding of marine systems. Additionally,

Biodiversity discovery and documentation has been transformed in the 21st century by molecular data to facilitate species identification, delineate species. And more recently, the development of high throughput sequencing technologies is used to quantify diversity and uncover large scale patterns of biodiversity organization. The wide adoption of DNA barcoding (where homologous DNA sequences are sequenced across the main domains of the Eukaryotes) has unraveled the pervasiveness of cryptic species, invigorating taxonomy and the description of new species, and enabled collaborative research programs such as the Census of Marine Life. Unraveling these cryptic species has also contributed to the recognition that species have smaller ranges than their taxonomic identity had implied making them potentially more susceptible to extinction or overexploitation. Despite early criticisms and now well-understood limitations, DNA barcoding has been suggested as a possible solution to the widening gap between the increased need to document biodiversity while taxonomic expertise is dwindling. Thus, large databases of DNA barcodes have been collaboratively assembled and made pubicly available. By matching DNA barcodes of unidentified species to a reference public database of accurately identified and vouchered specimens, DNA barcoding can greatly speed up species identification: sequences with no match make it easier to flag specimens that warrant further taxonomic attention.

Recent developments in high-throughput sequencing technologies are giving new applications to these DNA barcode databases [@Leray2016]. Metabarcoding has been developed to sequence in parallel the barcoding gene across the species included in environmental samples such as plankton tows, fish guts, or artificial structure scrappings [@Leray2015]. With decreasing sequencing costs, and the prospect of automating sample collection, these technologies will soon be used routinely for ecosystem and biodiversity monitoring [@Bohan2017]. DNA barcoding provides the possibility to link these newly generated data with historical natural history knowledge about species and their distributions. However, the insights we will gain from these approaches are only going to be as good as the level of completeness and accuracy of the information included in reference databases. Here, we mostly focus on museum specimens because we can validate their identification, and a fraction of these samples (if they have not been preserved in formalin) can be used to obtain DNA from them to complete existing databases of DNA barcodes.

@@

It has never been easier and cost effective to survey comprehensively the extant of marine biodiversity.
